---
title: "Collaborative Checklist"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Collaborative Checklist}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>",
 eval = FALSE
)
```

This vignette demonstrates how to build a collaborative checklist using autoedit's
`checklist_ui()` and `checklist_server()` functions. Unlike the textarea module,
the checklist provides a good collaborative experience because each action (add,
toggle, delete) is discrete - there's no cursor position to preserve.
## Why checklists work well for collaboration

Collaborative text editing is challenging because users have a cursor position
that must be preserved when remote changes arrive. Standard HTML textareas
can't do this, leading to a frustrating experience where your cursor jumps
around.

Checklists avoid this problem entirely:

- **Adding an item**: Click a button, type text, submit. The list re-renders
  with the new item - no cursor to lose.
- **Toggling an item**: Click a checkbox. The list re-renders with the updated
  state.
- **Deleting an item**: Click a delete button. The list re-renders without the
  item.

Each action is atomic and complete. When the UI re-renders after a remote
change, you're not in the middle of anything that gets disrupted.

Automerge handles concurrent operations correctly:

- Two users add items simultaneously: Both items appear (order may vary)
- Two users toggle the same item: Last write wins (acceptable for a boolean)
- One user deletes while another toggles: Delete wins (item is gone)

## Basic example

```{r basic}
library(shiny)
library(autoedit)

ui <- fluidPage(
  h3("Shared Shopping List"),
  checklist_ui("shopping")
)

server <- function(input, output, session) {
  checklist_server(
    "shopping",
    doc_id = "shopping-list",
    initial_items = c("Milk", "Bread", "Eggs")
  )
}

shinyApp(ui, server)
```

Open this app in two browser windows and try:

1. Adding items in both windows
2. Checking items off in one window and watching the other update
3. Deleting items

## Complete application: Team task board

This example shows a more complete application with multiple checklists for
different categories, plus a summary of progress:

```{r task-board}
library(shiny)
library(bslib)
library(autoedit)

ui <- page_fillable(
  padding = "1rem",
  h2("Team Task Board"),
  p(class = "text-muted", "Session: ", textOutput("session_id", inline = TRUE)),

  layout_columns(
    col_widths = c(4, 4, 4),

    card(
      card_header(class = "bg-danger text-white", "To Do"),
      card_body(checklist_ui("todo"))
    ),

    card(
      card_header(class = "bg-warning", "In Progress"),
      card_body(checklist_ui("in_progress"))
    ),

    card(
      card_header(class = "bg-success text-white", "Done"),
      card_body(checklist_ui("done"))
    )
  ),

  card(
    card_header("Progress Summary"),
    card_body(verbatimTextOutput("summary"))
  )
)

server <- function(input, output, session) {
  output$session_id <- renderText(substr(session$token, 1, 6))

  todo_items <- checklist_server(
    "todo",
    doc_id = "board-todo",
    initial_items = c("Design new feature", "Write tests", "Update docs")
  )

  in_progress_items <- checklist_server(
    "in_progress",
    doc_id = "board-in-progress",
    initial_items = c("Code review PR #42")
  )

  done_items <- checklist_server(
    "done",
    doc_id = "board-done",
    initial_items = c("Fix login bug", "Deploy v1.2")
  )

  output$summary <- renderPrint({
    todo <- todo_items()
    in_prog <- in_progress_items()
    done <- done_items()

    cat("To Do:       ", nrow(todo), "items",
        "(", sum(todo$done), "checked )\n")
    cat("In Progress: ", nrow(in_prog), "items",
        "(", sum(in_prog$done), "checked )\n")
    cat("Done:        ", nrow(done), "items",
        "(", sum(done$done), "checked )\n")
    cat("---\n")
    cat("Total:       ", nrow(todo) + nrow(in_prog) + nrow(done), "items\n")
  })
}

shinyApp(ui, server)
```

## Document IDs

Each `checklist_server()` call takes a `doc_id` parameter that identifies the
collaborative document. All sessions using the same `doc_id` will synchronize
together:

```{r doc-id-example}
# These will sync together (same doc_id)
checklist_server("list1", doc_id = "shared-list")
checklist_server("list2", doc_id = "shared-list")

# This is independent (different doc_id)
checklist_server("list3", doc_id = "private-list")
```

## Return value

`checklist_server()` returns a reactive expression containing a data frame with
the current items:
```{r return-value}
items <- checklist_server("tasks", doc_id = "my-tasks")

# In a reactive context:
observe({
  df <- items()
  # df has columns: id, text, done
  completed <- sum(df$done)
  total <- nrow(df)
  message(sprintf("%d/%d tasks completed", completed, total))
})
```

## Comparison with textarea

| Feature | `checklist_ui()` | `textarea_ui()` |
|---------|------------------|-----------------|
| Good collaborative UX | Yes | No (cursor jumps) |
| Concurrent edits merge well | Yes | Yes |
| Structured data | Yes (items) | No (plain text) |
| Free-form text | No | Yes |
| Use case | Task lists, voting | Notes, documents |

Use checklists when you have discrete items. Use the textarea (or better, the
full `editor()`) when you need free-form text editing.

## Limitations

Like the textarea module, the checklist module:
- Only syncs within a single R process (no multi-process scaling)
- Stores documents in memory (no persistence across restarts)

For production deployments needing persistence and scaling, consider storing
checklist state in a database and using Shiny's reactive system for
synchronization, or adapting the `editor()` approach with a sync server.
