---
title: "Collaborative Checklist and Kanban"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Collaborative Checklist and Kanban}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>",
 eval = FALSE
)
```

This vignette demonstrates how to build collaborative task management interfaces
using autoedit. We cover two approaches: simple checklists and kanban boards.
Both provide a good collaborative experience because each action (add, toggle,
delete, move) is discrete - there's no cursor position to preserve.
## Why checklists work well for collaboration

Collaborative text editing is challenging because users have a cursor position
that must be preserved when remote changes arrive. Standard HTML textareas
can't do this, leading to a frustrating experience where your cursor jumps
around.

Checklists avoid this problem entirely:

- **Adding an item**: Click a button, type text, submit. The list re-renders
  with the new item - no cursor to lose.
- **Toggling an item**: Click a checkbox. The list re-renders with the updated
  state.
- **Deleting an item**: Click a delete button. The list re-renders without the
  item.

Each action is atomic and complete. When the UI re-renders after a remote
change, you're not in the middle of anything that gets disrupted.

Automerge handles concurrent operations correctly:

- Two users add items simultaneously: Both items appear (order may vary)
- Two users toggle the same item: Last write wins (acceptable for a boolean)
- One user deletes while another toggles: Delete wins (item is gone)

## Basic example

```{r basic}
library(shiny)
library(autoedit)

ui <- fluidPage(
  h3("Shared Shopping List"),
  checklist_ui("shopping")
)

server <- function(input, output, session) {
  checklist_server(
    "shopping",
    doc_id = "shopping-list",
    initial_items = c("Milk", "Bread", "Eggs")
  )
}

shinyApp(ui, server)
```

Open this app in two browser windows and try:

1. Adding items in both windows
2. Checking items off in one window and watching the other update
3. Deleting items

## Complete application: Team task board

This example shows a more complete application with multiple checklists for
different categories, plus a summary of progress:

```{r task-board}
library(shiny)
library(bslib)
library(autoedit)

ui <- page_fillable(
  padding = "1rem",
  h2("Team Task Board"),
  p(class = "text-muted", "Session: ", textOutput("session_id", inline = TRUE)),

  layout_columns(
    col_widths = c(4, 4, 4),

    card(
      card_header(class = "bg-danger text-white", "To Do"),
      card_body(checklist_ui("todo"))
    ),

    card(
      card_header(class = "bg-warning", "In Progress"),
      card_body(checklist_ui("in_progress"))
    ),

    card(
      card_header(class = "bg-success text-white", "Done"),
      card_body(checklist_ui("done"))
    )
  ),

  card(
    card_header("Progress Summary"),
    card_body(verbatimTextOutput("summary"))
  )
)

server <- function(input, output, session) {
  output$session_id <- renderText(substr(session$token, 1, 6))

  todo_items <- checklist_server(
    "todo",
    doc_id = "board-todo",
    initial_items = c("Design new feature", "Write tests", "Update docs")
  )

  in_progress_items <- checklist_server(
    "in_progress",
    doc_id = "board-in-progress",
    initial_items = c("Code review PR #42")
  )

  done_items <- checklist_server(
    "done",
    doc_id = "board-done",
    initial_items = c("Fix login bug", "Deploy v1.2")
  )

  output$summary <- renderPrint({
    todo <- todo_items()
    in_prog <- in_progress_items()
    done <- done_items()

    cat("To Do:       ", nrow(todo), "items",
        "(", sum(todo$done), "checked )\n")
    cat("In Progress: ", nrow(in_prog), "items",
        "(", sum(in_prog$done), "checked )\n")
    cat("Done:        ", nrow(done), "items",
        "(", sum(done$done), "checked )\n")
    cat("---\n")
    cat("Total:       ", nrow(todo) + nrow(in_prog) + nrow(done), "items\n")
  })
}

shinyApp(ui, server)
```

## Document IDs

Each `checklist_server()` call takes a `doc_id` parameter that identifies the
collaborative document. All sessions using the same `doc_id` will synchronize
together:

```{r doc-id-example}
# These will sync together (same doc_id)
checklist_server("list1", doc_id = "shared-list")
checklist_server("list2", doc_id = "shared-list")

# This is independent (different doc_id)
checklist_server("list3", doc_id = "private-list")
```

## Return value

`checklist_server()` returns a reactive expression containing a data frame with
the current items:
```{r return-value}
items <- checklist_server("tasks", doc_id = "my-tasks")

# In a reactive context:
observe({
  df <- items()
  # df has columns: id, text, done
  completed <- sum(df$done)
  total <- nrow(df)
  message(sprintf("%d/%d tasks completed", completed, total))
})
```

---

## Kanban board

The checklist approach works well for simple lists, but sometimes you need to
organize items into columns - a kanban board. The `kanban_ui()` and
`kanban_server()` functions provide this, using a single shared Automerge
document where each item has a `column` field.

### Why a single document?

With separate checklists (as in the task board example above), items can't move
between columns - they're in separate documents. A kanban board uses one
document with all items, so moving an item just changes its `column` field.
Automerge syncs this change efficiently.

### Basic kanban example

```{r kanban-basic}
library(shiny)
library(autoedit)

ui <- fluidPage(
  h3("Team Kanban"),
  kanban_ui("board")
)

server <- function(input, output, session) {
  kanban_server(
    "board",
    doc_id = "team-kanban",
    initial_items = list(
      todo = c("Design feature", "Write tests"),
      in_progress = c("Code review"),
      done = c("Deploy v1.0")
    )
  )
}

shinyApp(ui, server)
```

### Complete kanban application

```{r kanban-complete}
library(shiny)
library(bslib)
library(autoedit)

ui <- page_fillable(
  padding = "1rem",
  h2("Collaborative Kanban Board"),
  p(class = "text-muted", "Session: ", textOutput("session_id", inline = TRUE)),
  kanban_ui("board"),
  hr(),
  card(
    card_header("Progress Summary"),
    card_body(verbatimTextOutput("summary"))
  )o
)

server <- function(input, output, session) {
  output$session_id <- renderText(substr(session$token, 1, 6))

  items <- kanban_server(
    "board",
    doc_id = "team-kanban",
    initial_items = list(
      todo = c("Design new feature", "Write tests", "Update docs"),
      in_progress = c("Code review PR #42"),
      done = c("Fix login bug", "Deploy v1.2")
    )
  )

  output$summary <- renderPrint({
    df <- items()
    for (col in c("todo", "in_progress", "done")) {
      col_items <- df[df$column == col, ]
      cat(sprintf("%-12s %d items (%d done)\n",
                  paste0(col, ":"),
                  nrow(col_items),
                  sum(col_items$done)))
    }
  })
}

shinyApp(ui, server)
```

Open in two browser windows and try:

1. Moving items between columns with the arrow buttons
2. Adding new items (they appear in "To Do")
3. Checking items off
4. Deleting items

All changes sync instantly to other sessions.

### Custom columns

You can define your own columns:

```{r kanban-custom}
kanban_ui(
  "board",
  columns = c(
    backlog = "Backlog",
    ready = "Ready",
    doing = "Doing",
    review = "Review",
    done = "Done"
  ),
  column_colors = c(
    backlog = "#64748b",
    ready = "#8b5cf6",
    doing = "#3b82f6",
    review = "#f59e0b",
    done = "#22c55e"
  )
)

kanban_server(
  "board",
  columns = c(
    backlog = "Backlog",
    ready = "Ready",
    doing = "Doing",
    review = "Review",
    done = "Done"
  ),
  initial_items = list(
    backlog = c("Future feature A", "Future feature B"),
    ready = c("Approved task"),
    doing = c("Current work"),
    review = c("PR #123"),
    done = c("Shipped!")
  )
)
```

---

## Comparison

| Feature | `checklist_ui()` | `kanban_ui()` | `textarea_ui()` |
|---------|------------------|---------------|-----------------|
| Good collaborative UX | Yes | Yes | No (cursor jumps) |
| Move items between groups | No | Yes | N/A |
| Structured data | Yes | Yes | No |
| Use case | Simple lists | Workflow stages | Notes |

Use checklists for simple task lists. Use kanban when items need to move through
stages. Use the `editor()` for free-form text editing.

## Limitations

Like the textarea module, the checklist module:
- Only syncs within a single R process (no multi-process scaling)
- Stores documents in memory (no persistence across restarts)

For production deployments needing persistence and scaling, consider storing
checklist state in a database and using Shiny's reactive system for
synchronization, or adapting the `editor()` approach with a sync server.
